---
title: "06 Drawdown & Crash Risk Analysis"
author: "Kimberly Lin"
date: "`r Sys.Date()`"
output: html_document
---

## 1. Motivation
While volatility captures short-term fluctuations, it does not fully reflect
the severity of downside risk faced by investors.

Drawdowns measure cumulative losses from peak to trough and provide a more
intuitive representation of market crashes and recovery dynamics.

This section examines drawdown behavior and crash risk across major
cryptocurrencies

## 2. Setup
```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(here)

source(here("src/utils/return_calculations.R"))
```

## 3. Load Data & Compute Log Returns
```{r load_data, message=FALSE}
btc <- read_csv(here("data/processed/BTC_clean.csv"))
eth <- read_csv(here("data/processed/ETH_clean.csv"))
bnb <- read_csv(here("data/processed/BNB_clean.csv"))

btc_ret <- calc_log_returns(btc)
eth_ret <- calc_log_returns(eth)
bnb_ret <- calc_log_returns(bnb)
```

## 4. Drawdown Calculation
```{r drawdown_cal, message=FALSE}
btc_dd <- compute_drawdown(btc_ret)
eth_dd <- compute_drawdown(eth_ret)
bnb_dd <- compute_drawdown(bnb_ret)
```

## 5. Drawdown Visualization
```{r drawdown_vis, message=FALSE}
plot_drawdown <- function(df, asset) {
  ggplot(df, aes(x = timeOpen, y = drawdown)) +
    geom_area(fill = "firebrick", alpha = 0.6) +
    labs(
      title = paste(asset, "Historical Drawdowns"),
      x = "Date",
      y = "Drawdown (log scale)"
    ) +
    theme_minimal()
}

plot_drawdown(btc_dd, "BTC")
plot_drawdown(eth_dd, "ETH")
plot_drawdown(bnb_dd, "BNB")
```

## 6. Maximum Drawdown Comparison
```{r max_drawdown_summary, message=FALSE}
max_dd_summary <- function(df, asset) {
  tibble(
    asset = asset,
    max_drawdown = min(df$drawdown, na.rm = TRUE)
  )
}

bind_rows(
  max_dd_summary(btc_dd, "BTC"),
  max_dd_summary(eth_dd, "ETH"),
  max_dd_summary(bnb_dd, "BNB")
)
```

## 7. Crash Risk：Extreme single-day loss
```{r crash_risk, message=FALSE, warning=FALSE}
detect_crash <- function(df, threshold = -0.10) {
  df %>%
    mutate(
      crash = log_return <= threshold
    )
}

btc_crash <- detect_crash(btc_ret)
eth_crash <- detect_crash(eth_ret)
bnb_crash <- detect_crash(bnb_ret)
```

## 8. Crash Frequency Summary
```{r crash_freq_summary, message=FALSE, warning=FALSE}
crash_summary <- function(df, asset) {
  df %>%
    summarise(
      asset = asset,
      crash_days = sum(crash, na.rm = TRUE),
      total_days = n(),
      crash_ratio = crash_days / total_days
    )
}

bind_rows(
  crash_summary(btc_crash, "BTC"),
  crash_summary(eth_crash, "ETH"),
  crash_summary(bnb_crash, "BNB")
)
```

```{r , message=FALSE, warning=FALSE}
```



```{r vol_regime, message=FALSE}
btc_ret <- calc_log_returns(btc)

btc_vol <- btc_ret %>%
  mutate(
    rolling_vol = zoo::rollapply(
      log_return,
      width = 30,
      FUN = sd,
      fill = NA,
      align = "right"
    )
  )

q <- quantile(btc_vol$rolling_vol, probs = c(0.33, 0.66), na.rm = TRUE)

btc_regime <- btc_vol %>%
  mutate(
    regime = case_when(
      rolling_vol <= q[1] ~ "Low Volatility",
      rolling_vol <= q[2] ~ "Medium Volatility",
      TRUE ~ "High Volatility"
    )
  )

```

## 9. Crash vs Volatility Regime
Do crash events occur disproportionately during high-volatility regimes?
```{r merge_regime_crash}
btc_regime_crash <- btc_regime %>%
  left_join(
    btc_crash %>% select(timeOpen, crash),
    by = "timeOpen"
  )
```
```{r crash_by_regime, message=FALSE, warning=FALSE}
btc_regime_crash %>%
  group_by(regime) %>%
  summarise(
    crash_days = sum(crash, na.rm = TRUE),
    total_days = n(),
    crash_ratio = crash_days / total_days
  )
```
Although high-volatility regimes account for fewer total days, they
contribute disproportionately to crash events.

```{r crash_ratio_plot, fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
btc_regime_crash %>%
  group_by(regime) %>%
  summarise(crash_ratio = mean(crash, na.rm = TRUE)) %>%
  ggplot(aes(x = regime, y = crash_ratio, fill = regime)) +
  geom_col() +
  labs(
    title = "Crash Probability by Volatility Regime (BTC)",
    x = "Volatility Regime",
    y = "Crash Probability"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r crash_overlay_plot, fig.height=5, fig.width=9, message=FALSE, warning=FALSE}
ggplot(btc_regime, aes(x = timeOpen, y = rolling_vol)) +
  geom_line(color = "steelblue") +
  geom_point(
    data = btc_regime_crash %>% filter(crash),
    aes(x = timeOpen, y = rolling_vol),
    color = "firebrick",
    size = 1.5,
    alpha = 0.7
  ) +
  labs(
    title = "Crash Events Overlayed on Rolling Volatility (BTC)",
    x = "Date",
    y = "30-Day Rolling Volatility"
  ) +
  theme_minimal()
```

Crash events cluster around periods of elevated volatility,
supporting regime-based risk modeling.

These results indicate that volatility regimes are not merely descriptive,
but carry meaningful predictive information regarding downside risk.

# 10. Key Insights
- Cryptocurrency markets experience **severe and prolonged drawdowns.**
- High-volatility regimes are strongly associated with crash events.
- BTC exhibits smaller maximum drawdowns relative to smaller-cap assets,
indicating partial risk mitigation from market dominance.
- Volatility alone understates downside risk; drawdown-based measures
provide essential insights for portfolio construction.