---
title: "08 Market Capitalization & Risk Structure Analysis"
author: "Kimberly Lin"
date: "`r Sys.Date()`"
output: html_document
---

## 1. Motivation
Market capitalization is often used as a proxy for liquidity, maturity,
and systemic importance in cryptocurrency markets.

Larger-cap assets are generally perceived as more stable, while smaller-cap
assets tend to exhibit higher volatility and crash risk.

This section investigates whether market capitalization is systematically
related to risk characteristics, including volatility, drawdowns, and tail risk.

## 2. Setup
```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(here)

source(here("src/utils/return_calculations.R"))
```

## 3. Load Data
```{r load_data, message=FALSE, warning=FALSE}
btc <- read_csv(here("data/processed/BTC_clean.csv"))
eth <- read_csv(here("data/processed/ETH_clean.csv"))
bnb <- read_csv(here("data/processed/BNB_clean.csv"))
```


## 4. Market Cap Overview
### 4-1. Market Cap Distribution
```{r market_cap_dist, message=FALSE}
bind_rows(
  btc %>% mutate(asset = "BTC"),
  eth %>% mutate(asset = "ETH"),
  bnb %>% mutate(asset = "BNB")
) %>%
  ggplot(aes(x = marketCap, fill = asset)) +
  geom_histogram(bins = 50, alpha = 0.6, position = "identity") +
  scale_x_log10() +
  labs(
    title = "Market Capitalization Distribution (Log Scale)",
    x = "Market Cap (log scale)",
    y = "Frequency"
  ) +
  theme_minimal()
```

## 5. Market Cap vs Volatility
### 5-1. Compute Volatility Summary
```{r cap_vs_vol, message=FALSE}
compute_vol_summary <- function(df, asset) {
  df %>%
    calc_log_returns() %>%
    summarise(
      asset = asset,
      avg_vol = sd(log_return, na.rm = TRUE),
      median_cap = median(marketCap, na.rm = TRUE)
    )
}

vol_cap_summary <- bind_rows(
  compute_vol_summary(btc, "BTC"),
  compute_vol_summary(eth, "ETH"),
  compute_vol_summary(bnb, "BNB")
)

vol_cap_summary
```

### 5-2. Visualization
```{r vol_cap_plot, message=FALSE}
ggplot(vol_cap_summary,
       aes(x = median_cap, y = avg_vol, label = asset)) +
  geom_point(size = 3, color = "steelblue") +
  geom_text(nudge_y = 0.001) +
  scale_x_log10() +
  labs(
    title = "Market Capitalization vs Volatility",
    x = "Median Market Cap (log scale)",
    y = "Return Volatility"
  ) +
  theme_minimal()
```

## 6. Market Cap vs Drawdown Severity

```{r cap_drawdown, message=FALSE}
compute_max_dd <- function(df, asset) {
  dd <- compute_drawdown(calc_log_returns(df))
  
  tibble(
    asset = asset,
    max_drawdown = min(dd$drawdown, na.rm = TRUE),
    median_cap = median(df$marketCap, na.rm = TRUE)
  )
}

dd_cap_summary <- bind_rows(
  compute_max_dd(btc, "BTC"),
  compute_max_dd(eth, "ETH"),
  compute_max_dd(bnb, "BNB")
)

dd_cap_summary
```

```{r dd_cap_plot, message=FALSE}
ggplot(dd_cap_summary,
       aes(x = median_cap, y = max_drawdown, label = asset)) +
  geom_point(size = 3, color = "firebrick") +
  geom_text(nudge_y = 0.05) +
  scale_x_log10() +
  labs(
    title = "Market Capitalization vs Maximum Drawdown",
    x = "Median Market Cap (log scale)",
    y = "Maximum Drawdown"
  ) +
  theme_minimal()
```

## 7. Interpretation
- Larger market-cap assets tend to exhibit **lower volatility and shallower drawdowns**
- Smaller-cap assets display **higher tail risk and more extreme downside behavior**
- Market capitalization acts as a **structural risk factor**, not merely a descriptive statistic 

## 8. Key Insights
- Market capitalization is **negatively related to volatility and drawdown severity**
- BTC’s dominant market cap contributes to its **relative stability**
- Smaller-cap cryptocurrencies carry **disproportionate downside and tail risk**
- Market cap should be treated as a **core risk dimension** in crypto portfolio construction
